Clouds and Radiation in a mock-Walker Circulation

### 2019-10-23 Silvers_WalkerCell.tex
changed exp to expl
added subsidence fraction to table 2, minor textual revisions

\documentclass[11pt]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{epstopdf}
%\usepackage{pdflatex}
%\usepackage{epsfig}

@@

\textbf{Key Points}
\begin{itemize}
  \item{At 25 km, simulations with parameterized shallow- and deep-convection exhibit asymmetries in the circulation, 
  precipitation, and humidity fields resulting in a steady-state overturning circulation that is not centered on the SST
  maximum.}
  \item{The longwave cloud-radiative effect is a fundamental factor in establishing these asymmetries and leads to dramatic
  changes in the low-level clouds and boundary layer structure.}
  \item{Cloud resolving simulations result in stronger overturning, more condensate aloft, and a RH in the deep convective 
  region that is 30\% higher than for the coarser cases with parameterized convection.}
  \item{The mock-Walker Circulation can be thought of as RCE plus an imposed gradient in surface temperature.}
\end{itemize}

@@

\begin{table}
\begin{center}
\caption{A few specifications of the simulations used most heavily in this study.  In the Convection
column, `prm'  indicates that convection is parameterized and `exp' indicates convection is explicity.  
The precipitation values in parenthesis are for the experiments in which LWCRE is off.}
    \begin{tabular}{*{6}{c}}
    \hline
    \hline
    \\
 Name & Grid Spacing $(\mathrm{km})$ & Domain $ (\mathrm{km^2}) $& Length $(\mathrm{months}) $ & Convection & 
 $\overline{\rm{P}} (\mathrm{mm/day})$    \\ \hline
  P100L &  100          &   800 $\times$ 16000    &  60              & prm           &  4.1 (3.5)          \\ 
    \\
  P100 &  100                & 800 $\times$ 4000     & 60            & prm         & 3.9 (3.7)               \\  
    \\
  P25L &  25             & 200 $\times$ 16000      & 60             & prm          & 4.0 (3.8)                \\  
    \\
  P25  &  25             & 200 $\times$ 4000      & 60             & prm           & 3.8 (3.7)               \\  
    \\
 E25  &   25          & 200 $\times$ 4000      & 60             & exp          & 3.7 (3.5)               \\  
    \\
 E2   &   2          & 100 $\times$ 4000      & 6             & exp          & 3.1 (3.4)                \\ 
    \\
 E1   &   1          & 10 $\times$ 4000      & 6             & exp          & 3.3 (3.7)                \\  \hline

\begin{table}
\begin{center}
\caption{Domain mean precipitation ($\overline{\rm{P}}$) and outgoing longwave radiation 
($\overline{\rm{OLR}}$).
The values in parenthesis are for the experiments in which LWCRE is off.}
    \begin{tabular}{*{3}{c}}
    \hline
    \hline
    \\
 Name &   $\overline{\rm{P}} (\mathrm{mm/day})$ & $\overline{\rm{OLR}} (\rm{W \, m^-2})$   \\ \hline
  P100L   &  4.1 (3.5)   &  283.1 (286.9)          \\ 
    \\
  P100 &   3.9 (3.7)   &  283.2 (296.4)             \\  
    \\
  P25L &   4.0 (3.8)  &  281.2 (290.7)             \\  
    \\
  P25  &   3.8 (3.7)    & 282.9 (293.6)             \\  
    \\
 E25  & 3.7 (3.5)    &   271.9 (286.8)               \\  
    \\
 E2   &  3.1 (3.4)   &  266.2 (285.5)               \\ 
    \\
 E1   &  3.3 (3.7)   &  269.3 (289.2)               \\  \hline

@@

@@

Superposing the circulation and relative humidity illustrates the subsidence driven 
drying and the strong moistening that results from ascent over the region of maximum latent heat flux.  
Figure \ref{fig:rh_psi_P25vsE25} shows two distinct circulation cells with one below, and one above $500 \rm{hPa}$.  
Deep-convective activity dominates the 
region over the warm pool, shallow convection is common over a wider range of SSTs, and for the most part stratocumulus 
clouds are absent.  When the coupling between the circulation and clouds is broken, by turning off the longwave CRE, the atmospheric 
state is much more symmetric about the maximum SST.  We also find large differences among the experiments in the domain mean 
precipitation, condensate/clouds, and circulation (see table 1).  In particular the experiments with parameterized convection are 
much more erratic.

@@

%The abnormality, or asymmetry, of the P25 simulation was the first surprising feature noticed by the authors.
%This asymmetry is also present on larger domains and at a resolution of $100 \, \rm{km}$.  For some of the other cases observed, such
%as P100 (on the small domain), the asymmetry can be described as an aversion of the precipitation maxima to reside over the 
%SST maximum while in other cases, such as the P25 case shown in \ref{fig:rh_psi_P25vsE25}, a strong ($1 \, \rm{m s^-1}$) domain
%mean shear develops which shifts the precipitation and circulation off center for years at a time.

@@

@@

Superposing the circulation and relative humidity illustrates the subsidence driven 
drying and the strong moistening that results from ascent over the region of maximum latent heat flux.

@@

An interesting fact that emerges from the domain mean values of precipitation (see Table 1) 
is that the sign of the response to LWCRE is not the same across the different grid-spacings.
When clouds are not allowed to interact with the longwave radiation the atmosphere can cool 
much more efficiently, as evidenced by larger values of OLR for all experiments when the LWCRE is
off.  Typically, the atmospheric cooling can be thought of as a proxy for the mean precipitation 
because the cooling is primarily offset by condensational heating.  This is clearly not the case 
for the E25, P25, and P100 cases for which the domain mean precipitation rates decrease despite
an increased amount of atmospheric cooling.  The implication is that the requisite atmospheric 
heating is coming from a process other than condensation.  The solution lies in the energy 
budget of the surface.  A fixed SST means that in all cases the upward flux of longwave energy
is constant, the upward flux of sensible heat flux will be mostly fixed (barring variations in surface
wind), and changes in downward shortwave flux will not be able to warm the surface.  However, 
the dramatic decrease of low-level clouds for the E25, P25, and P100 experiments when LWCRE is
off strongly influences the net flux of longwave energy at the surface (does the downwelling longwave
flux change becuase there are less clouds or because they are invisible to radiation?).  Under 
natural conditions, the surface is warmed by a downward flux of longwave radiation due to the clouds
above.  When this source of atmospheric energy loss is removed in the LWCRE off experiments 
the upwelling longwave flux of radiation plays a larger role in warming the atmosphere that more 
than compensates for the increased OLR.  Thus, with an increase in atmospheric warming on 
the order of $20 \, \rm{W \,m^{-2}}$ when then LWCRE is off, there is less need for condensational heating and the 
mean precipitation rate actually decreases.  The large decrease of low-level clouds also 
leads to a large increase of downward shortwave radiation.  The only slightly increases the
fluxes of reflected shortwave radiation (about $2 \, \rm{W \,m^{-2}}$) but also contributes to 
heating the atmosphere.
Turning the LW CRE off for the E1 and E2 experiments results in more precipitation while, 
for the E25, P25, and P100 
experiments, less precipitation results from turning the LW CRE off.  
The primary method by which the longwave cloud 
radiative effect influences the atmosphere is by heating the atmosphere in the region between the clouds and the surface.  
Larger values of ice condensate and upper-level cloud fraction such as we see for the E2 and E1 experiment 
therefore imply a larger LW CRE relative to the E25, P25, and P100 experiments in which there are fewer clouds 
aloft that can emit downwelling longwave radiation.  When the warming effect of the upper level clouds 
in the E1 and E2 experiments is removed in the LWCRE off experiments the energy balance of the atmosphere 
must be maintained through an increase of latent heating and subsequent increase of precipitation.  This is 
reflected in Table 1.  The decrease of precipitation for LWCRE off experiments for the P25 and P100 experiments
is harder to explain because there is much less upper level cloud to influence the mid-tropospheric heating.  
However, the low-level clouds of the P25 and P100 experiments are dramatically reduced when the LWCRE is off.  For 
GCM-like experiments with parameterized convection, the total precipitation is composed of one part from the 
convective parameterization, and one part from the large-scale cloud scheme.  Figure \ref{fig:conv_vs_ls} shows 
that when the LWCRE is turned off for the P25 and P100 experiments, the decrease in precipitation comes from 
a reduction in the large-scale cloud scheme precipitation and not from a decrease in the convective precipitation.       
This implies there will be less atmospheric heating from 
longwave radiation and a corresponding larger heating due to condensational heating which is consistent with higher values 
of domain mean precipitation for E25, P25, and P100 compared to E2 and E1.  Similarly, the E2 experiment has lower values 
of upper level condensate and higher values of mid-troposphere relative humidity compared to the E1 simulation, which ....??  
Water vapor is an efficient emitor of long-wave radiation and thus greatly facilitates the cooling of the troposphere.  
Popp and Silvers, 2017 showed that that there is dramatically less water vapor in the atmosphere and much less precipitation 
for LW CRE off experiments (see their figure 1).  Our results here for the E25, P25, and P100 are consistent with 
those of Popp and Silvers, 2017.

@@

>>>>

\documentclass[11pt]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
%\documentclass[draft]{agujournal2019}
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{epstopdf}
%\usepackage{pdflatex}
%\usepackage{epsfig}

@@

\textbf{Key Points: choose three}
\begin{itemize}
  \item{At 25 km, simulations with parameterized shallow- and deep-convection exhibit asymmetries in the circulation, 
  precipitation, and humidity fields resulting in a steady-state overturning circulation that is not centered on the SST
  maximum.}
  \item{Interactions between clouds and radiation combined with parameterized convection act to shift the precipitation 
  maximum away from the SST maximum.}
  \item{The longwave cloud-radiative effect is a fundamental factor in establishing these asymmetries and leads to dramatic
  changes in the low-level clouds and boundary layer structure.}
  \item{Cloud resolving simulations result in stronger overturning, more condensate aloft, and a RH in the deep convective 
  region that is 30\% higher than for the coarser cases with parameterized convection.}
  \item{The mock-Walker Circulation can be thought of as RCE plus an imposed gradient in surface temperature.}
\end{itemize}

@@

\begin{table}
\begin{center}
\caption{Specifications of the simulations used most heavily in this study.  In the Convection
column, `prm'  indicates that convection is parameterized and `expl' indicates explicit convection.}
    \begin{tabular}{*{5}{c}}
    \hline
    \hline
    \\
 Name & Grid Spacing $(\mathrm{km})$ & Domain $ (\mathrm{km^2}) $& Length $(\mathrm{months}) $ & Convection     \\ \hline
  P100L &  100          &   800 $\times$ 16000    &  60              & prm                   \\ 
    \\
  P100 &  100                & 800 $\times$ 4000     & 60            & prm                     \\  
    \\
  P25L &  25             & 200 $\times$ 16000      & 60             & prm                     \\  
    \\
  P25  &  25             & 200 $\times$ 4000      & 60             & prm                   \\  
    \\
 E25  &   25          & 200 $\times$ 4000      & 60             & expl                \\  
    \\
 E2   &   2          & 100 $\times$ 4000      & 6             & expl                   \\ 
    \\
 E1   &   1          & 10 $\times$ 4000      & 6             & expl                 \\  \hline

@@

\begin{table}
\begin{center}
\caption{Domain mean precipitation ($\overline{\rm{P}}$), outgoing longwave radiation 
($\overline{\rm{OLR}}$), and fraction of domain that is subsiding at the 532 hPa level.
Values in parenthesis correspond to experiments whith LWCRE is off.}
    \begin{tabular}{*{4}{c}}
    \hline
    \hline
    \\
 Name &   $\overline{\rm{P}} (\mathrm{mm/day})$ & $\overline{\rm{OLR}} (\rm{W \, m^-2})$ & SF   \\ \hline
  P100L   &  4.1 (3.5)   &  283.1 (286.9)    & 0.89 (0.61)      \\ 
    \\
  P100 &   3.9 (3.7)   &  283.2 (296.4)  & 0.73 (0.71)           \\  
    \\
  P25L &   4.0 (3.8)  &  281.2 (290.7)   & 0.78 (0.69)          \\  
    \\
  P25  &   3.8 (3.7)    & 282.9 (293.6)   & 0.80 (0.72)          \\  
    \\
 E25  & 3.7 (3.5)    &   271.9 (286.8)   & -            \\  
    \\
 E2   &  3.1 (3.4)   &  266.2 (285.5)    & 0.83 (0.77)           \\ 
    \\
 E1   &  3.3 (3.7)   &  269.3 (289.2)    & 0.78 (0.78)         \\  \hline

@@

Superposing the circulation and relative humidity illustrates the subsidence driven 
drying and the strong moistening that results from ascent over the region of maximum latent heat flux.  
Figure \ref{fig:rh_psi_P25vsE25} shows two distinct circulation cells with one below, and one above $500 \rm{hPa}$.  
Deep-convective activity dominates the 
region over the warm pool, shallow convection is common over a wider range of SSTs, and for the most part stratocumulus 
clouds are absent.  When the coupling between the circulation and clouds is broken, by turning off the longwave CRE, 
the atmospheric state is much more symmetric about the maximum SST.  We also find large differences among the 
experiments in the domain mean precipitation, condensate/clouds, and circulation (see Table 2).  In particular the 
experiments with parameterized convection are much more erratic.

@@

@@

Previous studies have shown that cloud radiative effect act to strengthen and contract an overturning 
(e.g. Popp and Silvers, 2017, and others (Muller and Held?)).  This was shown to be due to an increased low-level flux
of moist static energy into the convective regions.

@@

@@

An interesting fact that emerges from the domain mean values of precipitation (see Table 1) 
is that the sign of the response to LWCRE is not the same across the different grid-spacings.
When clouds are not allowed to interact with the longwave radiation the atmosphere can cool 
much more efficiently, as evidenced by larger values of OLR for all experiments when the LWCRE is
off.  Typically, the atmospheric cooling can be thought of as a proxy for the mean precipitation 
because the cooling is primarily offset by condensational heating.  This is clearly not the case 
for the E25, P25, and P100 cases for which the domain mean precipitation rates decrease despite
an increased amount of atmospheric cooling.  The implication is that the requisite atmospheric 
heating is coming from a process other than condensation.  The solution lies in the energy 
budget of the surface.  A fixed SST means that in all cases the upward flux of longwave energy
is constant, the upward flux of sensible heat flux will be mostly fixed (barring variations in surface
wind), and changes in downward shortwave flux will not be able to warm the surface.  However, 
the dramatic decrease of low-level clouds for the E25, P25, and P100 experiments when LWCRE is
off strongly influences the net flux of longwave energy at the surface (does the downwelling longwave
flux change becuase there are less clouds or because they are invisible to radiation?).  Under 
natural conditions, the surface is warmed by a downward flux of longwave radiation due to the clouds
above.  When this source of atmospheric energy loss is removed in the LWCRE off experiments 
the upwelling longwave flux of radiation plays a larger role in warming the atmosphere that more 
than compensates for the increased OLR.  Thus, with an increase in atmospheric warming on 
the order of $20 \, \rm{W \,m^{-2}}$ when then LWCRE is off, there is less need for condensational heating and the 
mean precipitation rate actually decreases.  The large decrease of low-level clouds also 
leads to a large increase of downward shortwave radiation.  This only slightly increases the
fluxes of reflected shortwave radiation (about $2 \, \rm{W \,m^{-2}}$) but does contribute to 
heating the atmosphere.

Turning the LW CRE off for the E1 and E2 experiments results in more precipitation while, 
for the E25, P25, and P100 experiments, less precipitation results from turning the LW CRE off.  
The primary method by which the longwave cloud radiative effect influences the atmosphere is by 
heating the atmosphere in the region between the clouds and the surface.  
Larger values of ice condensate and upper-level cloud fraction such as we see for the E2 and E1 experiment 
therefore imply a larger LW CRE relative to the E25, P25, and P100 experiments in which there are fewer clouds 
aloft that can emit downwelling longwave radiation.  When the warming effect of the upper level clouds 
in the E1 and E2 experiments is removed in the LWCRE off experiments the energy balance of the atmosphere 
must be maintained through an increase of latent heating and subsequent increase of precipitation.  This is 
reflected in Table 1.  The decrease of precipitation for LWCRE off experiments for the P25 and P100 experiments
is harder to explain because there is much less upper level cloud to influence the mid-tropospheric heating.  
However, the low-level clouds of the P25 and P100 experiments are dramatically reduced when the LWCRE is off.

For GCM-like experiments with parameterized convection, the total precipitation is composed of one part from the 
convective parameterization, and one part from the large-scale cloud scheme.  Figure \ref{fig:conv_vs_ls} shows 
that when the LWCRE is turned off for the P25 and P100 experiments, the decrease in precipitation comes from 
a reduction in the large-scale cloud scheme precipitation and not from a decrease in the convective precipitation.       
This implies (is this connected to previous sentence?) there will be less atmospheric heating from 
longwave radiation (but doesn't LW just cool the atmosphere?) and a corresponding 
larger heating due to condensational heating which is consistent with higher values 
of domain mean precipitation for E25, P25, and P100 compared to E2 and E1.  Similarly, the E2 experiment has lower values 
of upper level condensate and higher values of mid-troposphere relative humidity compared to the E1 simulation, which ....??  
Water vapor is an efficient emitter of long-wave radiation and thus greatly facilitates the cooling of the troposphere.  
Popp and Silvers, 2017 showed that that there is dramatically less water vapor in the atmosphere and much less precipitation 
for LW CRE off experiments (see their figure 1b).  Our results here for the E25, P25, and P100 are consistent with 
those of Popp and Silvers, 2017 because the low-level condensate strongly decreases when the LWCRE is off.  This is not 
the case for E2 and E1.  Those experiments have very little low-level condensate to begin with and thus do not experience 
large changes in low-level condensate when the LWCRE is off.  They do however have a large increase of upper
level condensate for the LWCRE off experiments.   This is a large difference in the way that models with explicit convection 
respond to the LWCRE compared to more traditional large-scale models with parameterized convection.

@@

%\begin{thebibliography}{}
%
%\bibitem[{\textit{Bretherton et~al.}(2005)\textit{Bretherton, Blossey, and
%  Khairoutdinov}}]{Bretherton_etal_2005}
%Bretherton, C.~S., P.~N. Blossey, and M.~Khairoutdinov (2005), An
%  energy-balance analysis of deep convective self-aggregation above uniform
%  sst, \textit{J. Atmos. Sci.}, \textit{62}, 4273--4292.
%
%%\bibitem[{\textit{Jeevanjee and Romps}(2013)}]{Jeevanjee_Romps_2013}
%%Jeevanjee, N., and D.~M. Romps (2013), Convective self-aggregation, cold pools,
%%  and domain size, \textit{Geophys.\ Res.\ Lett.}, \textit{40}, 994--99,
%%  \doi{10.1002/grl.50204}.
%  
%\end{thebibliography}

%%% Acknowledgements
%\begin{acknowledgments}
%(Levi's acknowledgements)
%\end{acknowledgments}
